---
title: Practice for 'R for Biologists'
author: Liang
date: '2018-07-16'
slug: practice-for-r-for-biologists
categories:
  - R
tags:
  - R
---

# Part 1 Basic
R displays 7 digits by default. You can display more digits with options(). More than 15 digits could be unreliable. This is a **global option**; remains in effect until further notice
```{r}
options(digits = 15)
2/3; 2.1^3.1
```

built-in mathematical constants and functions
```{r}
2*pi
sin(2*pi)
 
# This is e
exp(1)               
 
sqrt(2)
```

cat() function concatenates strings, inside double quotes, and interpolated values of variables  

```{r}
x= 2.1
cat("x =", x, "\n")
cat("root2 =", sqrt(2), "\n", "root3 =", sqrt(3), "\n")
```

# Part 2 Help

```{r}
# To see a sequence of simple graphics pictures
demo(graphics)        
 
# To see a sequence of 3D graphics
#demo(persp)           
 
# To list available demonstrations in base package of R
#demo() 
```


```{r}
# To list all arithmetic operators in R
help(Arithmetic)      
 
# To get help on the log function
help(log)             
 
# To display R help on the plot() function
help(plot)            
 
# Same as help(plot)
?plot                 
 
# To see example of the usage of plot function
example(plot)         
 
# To list all R commands containing the string plot; could be too much info
help.search("plot")   
 
# Same as help.search("plot")
??plot   
```

: operator has higher predecence than the arithmetical operations.

```{r}
sequence_1 = 0 : 10 - 2
print(sequence_1)
 
sequence_2 = 0 : (10 - 2)
print(sequence_2)
```

```{r}
x = seq(from = 0, to = 10, by = 0.5)
square_root_x = sqrt(x)
 
plot(x, square_root_x)
```


# Part 3 Graphics

```{r}
x = seq(from =0, to =10, by=0.5)
square_root_x = sqrt(x)
plot(x, square_root_x)

# To pause between plots
par(ask = TRUE) 

# To add title
plot(x, square_root_x, main = "My First Plot")                         
 
# To add label to y-axis
plot(x, square_root_x, main = "My First Plot", ylab = "sqrt(x)")        
 
# To set limits of y-axis
plot(x, square_root_x, main = "My First Plot", ylab = "sqrt(x)", ylim = c(0, 10))  
 

# add color
plot(x, square_root_x, main = "my fisrt plot", ylab = "sqrt(x)", ylim = c(0,10), col = "blue")

# To over strike with both plotting characters, pch, and connecting lines
plot(x, square_root_x, main = "My First Plot", ylab = "sqrt(x)", ylim = c(0, 10),col = "blue", type = "o")  

# To set the line type
plot(x, square_root_x, main = "My First Plot", ylab = "sqrt(x)", ylim = c(0, 10), col = "blue", type = "o", lty = "dotted") 

# To set plotting character 
plot(x, square_root_x, main = "My First Plot", ylab = "sqrt(x)", ylim = c(0, 10), col = "blue", type = "o", lty = "dotted", pch = 22)   
```


plot() is a high-level plotting function that opens a new plotting window, using low-level functions, one can add to the open plot,some of the commonly used adding functions are: points(), lines(), abline(), legend(), text() 

```{r}
t = seq(from = 0, to = 10, by = 0.5)
        
# To pause between plots
#par(ask = TRUE)                                           
    
plot(t, sqrt(2 * t), main = "Adding to Plot", sub = "A function and its inverse", xlab = "", ylab = "", xlim = c(0, 10), ylim = c(0, 10), pch = 16, type = "o") 
 
# To add the inverse function
points(t, 0.5 * t * t, col = "red" , pch = 16, type = "o")    
 
# To add the line with intercept = 0, slope = 1
abline(0, 1, col = "gray", lwd = 3, lty = "dashed")      
 
# To add text
text(7, 7.3, "45-degree line", srt = 45)                 
legend("bottomright", legend = c("funtion", "inverse"), pch = 16, col = c("black", "red"))
 
# One can add graphs of mathematical functions using, for example:
x = seq(from =0, to =10, by=0.5)
lines(curve(sin(x) + cos(x) + 5,  add = TRUE))
```

curve() is also a high-level mathematical plotting function that opens a new plotting window, Variable in the mathematical formula must be x, add more graphs by setting add = TRUE
```{r}
curve(sin(x), xlim = c(-4*pi, 4*pi), ylim = c(-2, 2), col = "red", ylab = "")
 
# To plot horizontal line using h
abline(h=0, lty= "dotted")    
 
# To mark the origin      
points(0, 0, pch = 3)                
 
curve(cos(x),  col = "blue", add = TRUE) 
curve(sin(x) + cos(x),  col = "purple", lwd = 3, add = TRUE)
 
title("sin(x) + cos(x)")
 
legend("bottomright", legend = c("sin(x)", "cos(x)", "sin(x) + cos(x)"), 
        lty = 1, col = c("red", "blue", "purple"))
```

par() command can set the multi plot environment. mfcol = c(nr,nc) partitions the graphic window as a matrix of nr rows and nc columns, the plots are then drawn in columns. mfrow = c(nr,nc) partitions the graphic window as a matrix of nr rows and nc columns, the plots are then drawn in rows. You can get fancy partitions of the plotting window with layout() function

```{r}
# Prepare for 2x2 plots, to be filled by rows
par(mfrow = c(2, 2))          
 
t = seq(0, 10, 0.2)
plot(t, sin(t)) 
curve(sinh(x), -5, 5)
curve(tan(x), n = 500, -5, 5)
curve(round(x), n = 500, -5, 5) 
 
# Return to single plot
par(mfrow = c(1, 1))          
 
curve(sin(x), -5, 5) 
```


# Part 4 Curve Fitting

Linear Model function lm(y~x) computes the least square fit line to data points; y as a function of x. Type ?formula for help with legal formulas. lm() can also be used to fit nonlinear models where the parameters enter into model linearly

```{r}
# fit a line to two data points
x2 = c(1,2)
y2 = c(1.5,2.5)
plot(x2,y2, main = "Unique line through 2 points", xlim = c(0,4), ylim = c(0, 4))

# compute the least square line 

ls_fit_line = lm(y2~x2)
print(ls_fit_line)

# plot the least square line stored in ls_fit_line
abline(ls_fit_line, col = "red")

# plot the vector contain the intercept and slope 
print(coef(ls_fit_line))

# more information
print(summary(ls_fit_line))
```

```{r}
# Compute least-square line fit for 3 data points
x3 = c(1, 2, 3)
y3 = c(1.5, 2.5, 2.8)
 
plot(x3, y3, main="Best-fit line through 3 points", xlim=c(0, 4), ylim = c(0, 4)) 
fit3 = lm(y3 ~ x3)
abline(fit3, col = "red")
```

```{r}
print(fit3)
print(summary(fit3))
```
```{r}
# Compute the parabola through 3 data points
 
plot(x3, y3, main="Unique parabola through 3 points", xlim=c(0, 6), ylim = c(0, 6)) 
fit_parabola3 = lm(y3 ~ x3 + I(x3^2))
print(fit_parabola3 )
curve(coef(fit_parabola3)[1] + coef(fit_parabola3)[2]* x + coef(fit_parabola3)[3]*x^2, 0, 6, col = "red", add=TRUE)
```

```{r}
# Compute a best-fit parabola for 4 data points
 
x4 = c(1, 2, 3, 4)
y4 = c(1.5, 2.5, 2.8, 1.7)
 
plot(x4, y4, main="Best-fit parabola through 4 points", xlim=c(0, 6), ylim = c(0, 6)) 
fit_parabola4 = lm(y4 ~ x4 + I(x4^2))
curve(coef(fit_parabola4)[1] + coef(fit_parabola4)[2]* x + coef(fit_parabola4)[3]*x^2, 0, 6, col = "red", add=TRUE)
 
# Add the predicted points
points(x4, predict(fit_parabola4), col = "green") 

print(fit_parabola4)
```







