---
title: Notes From Advanced R Part One
author: Liang
date: '2018-12-02'
slug: notes-from-advanced-r
categories:
  - R
tags: []
lastmod: '2018-12-02T21:44:12-05:00'
keywords: []
description: ''
comment: no
toc: no
autoCollapseToc: no
postMetaInFooter: no
hiddenFromHomePage: no
contentCopyright: no
reward: no
mathjax: yes
mathjaxEnableSingleDollar: yes
mathjaxEnableAutoNumber: yes
hideHeaderAndFooter: no
flowchartDiagrams:
  enable: no
  options: ''
sequenceDiagrams:
  enable: no
  options: ''
---

# 1. Data structure 

The three properties of a vector are type, length, and attributes.

All objects can have arbitrary additional attributes

Attributes can be accessed individually with `attr()` or all at once (as a list) with `attributes()`.

```{r}
y <- 1:10
attr(y, "my_attribute") <- "This is a vector"
attr(y, "my_attribute")
```


# 2. Subsetting

`outer()` function applies a function to two arrays.

```{r}
x <- c(1, 2.3, 2, 3, 4, 8, 12, 43)
y<- c(2, 4)
outer(x, y, "+")
```

Multiply array x elements with array y elements:
```{r}
x %o% y  #equal to outer(x,y,"*")
```

```{r}
vals <- outer(1:5, 1:5, FUN = "paste", sep = ",")
vals
```

`[` will simplify the results to the lowest possible dimensionality. add `drop = F`

```{r}
vals[,1]
vals[,1, drop= F]
```

Subsetting operators diffenent in S3 and S4 objects. In S4 object, `@` (equivalent to `$`), and `slot()` (equivalent to `[[`). `@` is more restrictive than `$` in that it will return an error if the slot does not exist.  

`upper.tri` returns a matrix of logicals the same size of a given matrix with entries TRUE

```{r}
x <- outer(1:5, 1:5, FUN = "*")
upper.tri(x)
```

`diag()` function extracts or replaces the diagonal of a matrix, or constructs a diagonal matrix.

```{r}
diag(3)
diag(10,3,4) # guess what?
```

**Factor: drops any unused levels.**
```{r}
z <- factor(c("a", "b"))
z[1]
z[1, drop=T]
```

`$` diffenent with `[[`

```{r}
var <- "cyl"
# Doesn't work - mtcars$var translated to mtcars[["var"]]
mtcars$var
# mtcars@var doesn't work
mtcars[[var]]
```

`$` does partial matching while `[[` doesn't

```{r}
x <- list(abc = 1)
x
x$a
x[["a"]]
```

```{r}
mod <- lm(mpg ~ wt, data = mtcars)
mod
```

`summary` is a generic function used to produce result summaries of the results of various model fitting functions. 

```{r}
summary(mod)
```

Subsetting with nothing can be useful in conjunction with assignment because it will preserve the original object class and structure.

```{r}
mtcars[] <- lapply(mtcars, as.integer)
mtcars
mtcars <- lapply(mtcars, as.integer)
mtcars
```

# 3. Vocabulary

`on.exit()` can be used in the function to perform some side effect. For example, in addition to returning a value, the following function uses on.exit() to also print two messages.

```{r}
myfun = function(x){
        on.exit(print("first"))
        on.exit(print("second"), add = TRUE)
        return(x)
}
myfun(2)
```

remove add=TRUE from the second `on.exit()` usage.

```{r}
fun = function(x){
        on.exit(print("first"))
        on.exit(print("second"))
        return(x)
}
fun(2)
```

```{r}
fun = function(x){
        return(print("first"))
        return(print("second"))
        return(x)
}
fun(2)
```
`invisible` Return a (temporarily) invisible copy of an object.
```{r}
f1 <- function(x) x
f2 <- function(x) invisible(x)
f1(1)  # prints
f2(1)  # does not
```

```{r}
f1 <- function(x){
  if( x > 0 ){
     invisible("bigger than 0")
  }else{
     return("negative number")
  }
  "something went wrong"
}
f1(1)
```

```{r}
f2 <- function(x){
  if( x > 0 ){
     return(invisible("bigger than 0"))
  }else{
     return("negative number")
  }
}

f2(1) # not visible
```

`invisible` not print its return value but pass on the value as usual:

```{r}
a <- f2(1)
a
```

force invisible to visible 

```{r}
f1 <- function() 1
f2 <- function() invisible(1)
f1()
f2()
(f2())
```


# 4. Style

Nothing New.

Commenting guidelines

```{r}
# Load data ---------------------------

# Plot data ---------------------------
```

# 5. Functions

R function have three parts: `body()`, `formals()`, `environment()`.

```{r}
f <- function(x) x^2
f

formals(f)
body(f)
environment(f)
```

The `body()`, `formals()`, `environment()` call of Primitive functions is NULL.

Primitive functions are in base package and they are more efficient

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
objs[1]
funs <- Filter(is.function, objs)
funs[1]
```

> is.function estimate an object is a function or not
> Filter is powerful than filter in dplyr, Filter extracts the elements of a vector for which a predicate (logical) function gives true
> mget: Return the Value of a Named Object

four basic principles behind R’s implementation of lexical scoping:

- name masking
- functions vs. variables
- a fresh start
- dynamic lookup

```{r}
x <- 1
h <- function() {
  y <- 2
  i <- function() {
    z <- 3
    c(x, y, z)
  }
  i()
}
h()
rm(x, h) # equal i <- c(1, 2, 3)
```

```{r}
j <- function(x) {
  y <- 2
  function() {
    c(x, y)
  }
}
k <- j(1)
k()
rm(j, k)
```
tweak rules, R will ignore objects that are not functions while it is searching both a name of variables and functions
```{r}
n <- function(x) x / 2
o <- function() {
  n <- 10
  n(n)
}
o()
```

`codetools::findGlobals()` can be used to lists all the external dependencies of a function

```{r}
f <- function() x + 1
codetools::findGlobals(f)
```

```, the backtick, can be used to refer a functions or variables that have otherwise reserved or illegal names:

```{r}
for (i in 1:2) print(i)
`for`(i, 1:2, print(i))
```

```{r}
{print(1); print(2); print(3)}
`{`(print(1), print(2), print(3))
```

```{r}
x <- list(1:3, 4:9, 10:12)
sapply(x, "[", 2)
# equivalent to
sapply(x, function(x) x[2])
```

Arguments are matched first by exact name (perfect matching), then by prefix matching, and finally by position.

`stop`: stops execution of the current expression and executes an error action.

```{r}
iter <- 12
try(if(iter > 10) stop("too many iterations"))

tst1 <- function(...) stop("dummy error")
try(tst1(1:10, long, calling, expression))

tst2 <- function(...) stop("dummy error", call. = FALSE)
try(tst2(1:10, longcalling, expression, but.not.seen.in.Error))
```


```{r}
add <- function(x) {
  function(y) x + y
}
adders <- lapply(1:10, add) # make a list of function; x is from 1 to 10
adders[[1]](10)  # get the first function and supply y with 10
```

Default arguments are evaluated inside the function. 
```{r}
rm(list = ls()) # remove all variable
f <- function(x = ls()) {
  a <- 1
  x
}
# ls() evaluated inside f:
f()
# ls() evaluated in global environment 
f(ls())
```

`...` will match any arguments not otherwise matched, and can be easily passed on to other functions.

```{r}
f <- function(...) {
  names(list(...))
}
f(a = 1, b = 2)
```

Using `...` comes at a price — any misspelled arguments will not raise an error, and any arguments after ... must be fully named. This makes it easy for typos to go unnoticed:

```{r}
sum(1, 2, NA, na.mr = TRUE)
sum(1, 2, NA, na.rm = TRUE)
```

The below is the same, notice the `` function

```{r}
`second<-` <- function(x, value) {
  x[2] <- value
  x
}
x <- 1:10
second(x) <- 5L
x
```

```{r}
second <- function(x, value) {
  x[2] <- value
  x
}
x <- 1:10
second(x) <- 5L
x
```
using pryr::address() to find the memory address, find the address is different
```{r}
library(pryr)
x <- 1:10
address(x)
second(x) <- 6L
address(x)
```

but the default subset is not change the address
```{r}
x <- 1:10
address(x)

x[2] <- 7L
address(x)
```

```{r}
x <- 1:10
`modify<-` <- function(x, position, value) {
  x[position] <- value
  x
}
modify(x, 1) <- 10
x
x <- 1:10
x <- `modify<-`(x, 1, 10)
x
x <- 1:10
# modify(get("x"), 1) <- 10 # do not work because this the same with 
# get("x") <- `modify<-`(get("x"), 1, 10)
```

































