---
title: "Meachine learning in R"
author: "KJY"
date: '2022-01-01'
slug: mechine_learning_in_R
categories: R
tags: []
lastmod: '2022-01-01T10:29:22-05:00'
keywords: []
description: ''
comment: no
toc: no
autoCollapseToc: no
postMetaInFooter: no
hiddenFromHomePage: no
contentCopyright: no
reward: no
mathjax: yes
mathjaxEnableSingleDollar: yes
mathjaxEnableAutoNumber: yes
hideHeaderAndFooter: no
flowchartDiagrams:
  enable: no
  options: ''
sequenceDiagrams:
  enable: no
  options: ''
---

```{r}
library(tidyverse)
```

## k-means clustering

The k-means clustering algorithms aims at partitioning n observations into a fixed number of k clusters. The algorithm will find homogeneous clusters.


```{r eval=FALSE, include=FALSE}
stats::kmeans(x, centers = 3, nstart = 10)
```


where

x is a numeric data matrix
centers is the pre-defined number of clusters
the k-means algorithm has a random component and can be repeated nstart times to improve the returned model

```{r}
i <- grep("Length", names(iris))
x <- iris[, i]
cl <- kmeans(x, 3, nstart = 10)
plot(x, col = cl$cluster)
```

```{r}
sample_n(x, 5)
```

```{r}
cl$cluster
```


```{r}
cl
```


### How does k-means work

Initialisation: randomly assign class membership

```{r}
set.seed(12)
init <- sample(3, nrow(x), replace = TRUE)

init
plot(x, col = init)
```

Iteration:

Calculate the centre of each subgroup as the average position of all observations is that subgroup.
Each observation is then assigned to the group of its nearest centre.
It’s also possible to stop the algorithm after a certain number of iterations, or once the centres move less than a certain distance.

```{r}

centres <- sapply(1:3, function(i) colMeans(x[init == i, ], ))

centres

centres <- t(centres)

points(centres[, 1], centres[, 2], pch = 19, col = 1:3)

centres

rbind(centres, x)

tmp <- dist(rbind(centres, x))

ki <- apply(tmp, 1, which.min)
ki <- ki[-(1:3)]
```



```{r}
par(mfrow = c(1, 2))
plot(x, col = init)

# calculate mean value of each cluster now
centres <- sapply(1:3, function(i) colMeans(x[init == i, ], ))
centres <- t(centres)
points(centres[, 1], centres[, 2], pch = 19, col = 1:3)

#calculate Euclidean distance between each row in matrix
tmp <- dist(rbind(centres, x))
tmp <- as.matrix(tmp)[, 1:3]

ki <- apply(tmp, 1, which.min)
ki <- ki[-(1:3)]

plot(x, col = ki)
points(centres[, 1], centres[, 2], pch = 19, col = 1:3)
```


Termination: Repeat iteration until no point changes its cluster membership.

```{r}
cl$tot.withinss
```


```{r}
ks <- 1:5
tot_within_ss <- sapply(ks, function(k) {
    cl <- kmeans(x, k, nstart = 10)
    cl$tot.withinss
})
plot(ks, tot_within_ss, type = "b")
```






