---
title: Add p value and significant marker for ggplot based on ggpubr
author: Liang
date: '2018-06-07'
slug: add-p-value-and-significant-marker-for-ggplot-based-on-ggpubr
categories:
  - Bioinfo
tags:
  - Bioinfo
---

>The ‘ggpubr’ package provides some easy-to-use functions for creating and customizing ‘ggplot2’- based publication ready plots. -A. Kassambara.

[Reference](http://www.sthda.com/english/rpkgs/ggpubr)

# 1. Preparation

install the package

```{r eval=FALSE, message=FALSE, warning=FALSE}
install.packages("ggpubr")
```
or you can install the latest version form github 

```{r eval=FALSE}
if(!require(devtools)) install.packages("devtools") # if havn`t install devtools before, install it first
devtools::install_github("kassambara/ggpubr")
```
load package 

```{r}
library(ggpubr)
```

load data

```{r}
data("ToothGrowth")
head(ToothGrowth)
```

# 2. Comparing means in R 

Reference here:
http://www.sthda.com/english/wiki/comparing-means-in-r


a short summary here

Methods | R function | Describe
-------|------------|----------
T test | t.test() | two groups (parametric)
Wilcoxon test | wilcox.test | two groups (nonparametric)
ANOVA | anova() | multiple groups (parametric)
Kruskal-Wallis | kruskal.test() | multiple groups (nonparametric)

# 3. Two functions 

* compare_means()
* stat_compare_means()

usage

```
compare_means(formula, data, method="wilcox.test", paired=FALSE, 
              group.by=NULL, ref.group = NULL, ...)
```

>formula: a formula of the form x ~ group where x is a numeric variable giving the data values and group is a factor with one or multiple levels giving the corresponding groups 

>ref.group: a character string specifying the reference group. If specified, for a given grouping variable, each of the group levels will be compared to the reference group

```
stat_compare_means(mapping = NULL, 
                   comparisons = NULL,
                   hide.ns = FALSE,
                   label = NULL,
                   label.x = NULL,
                   label.y = NULL)
```

# 4. Compare two group 

```{r message=FALSE, warning=FALSE}
compare_means(len ~ supp, data = ToothGrowth) 
```

```{r}
p <- ggboxplot(ToothGrowth, x="supp",
               y = "len", color = "supp", add = "point")
# add p value
p + stat_compare_means() # the default method for comparision is wilcox.test() 
```

change comparision method

```{r}
p + stat_compare_means(method = "t.test")
```

add the p.signif and change the location
```{r}
p + stat_compare_means(aes(label = ..p.signif..),
                       label.x = 1.5, 
                       label.y = 40)
```

# 5. Compare two paired group 

```{r}
compare_means(len ~ supp, data = ToothGrowth, paired = TRUE)

ggpaired(ToothGrowth, x="supp", y="len",
         color="supp", line.color="gray",
         line.size=0.4, palette = "point") + 
    stat_compare_means(paired = TRUE)
```

# 6. multiple group comparision 

global test 

```{r}
compare_means(len ~ dose, data = ToothGrowth, method = "anova")
```

```{r}
ggboxplot(ToothGrowth, x = "dose", y = "len", color = "dose", add = "point") + 
    stat_compare_means(method = "anova")
```

you can specify the comparison group

```{r}
my_comparisons <- list(c("0.5","1"), c("1", "2"),
                       c("0.5", "2"))
ggboxplot(ToothGrowth, x="dose", y="len",
          color="dose", add = "point") +
    stat_compare_means(comparisons = my_comparisons) + #add comparision group p value 
    stat_compare_means(label.y = 50) # add global p value 
```

# 7. multiple group comparision with a reference group 

```{r}
compare_means(len ~ dose, data = ToothGrowth,
              ref.group = "0.5", 
              method = "t.test")
```

Visualize the result 

```{r}
ggboxplot(ToothGrowth, x="dose", y="len",
          color="dose", add = "point") + 
    stat_compare_means(method="anova", label.y=40) + # add global p value 
    stat_compare_means(label="p.signif", method="t.test",
                       ref.group = "0.5") # add signif marker
```

use all data as the base-mean do the paired comparison

```{r}
# Load myeloma data from GitHub
myeloma <- read.delim("https://raw.githubusercontent.com/kassambara/data/master/myeloma.txt")

compare_means(DEPDC1 ~ molecular_group,  data = myeloma,
              ref.group = ".all.", method = "t.test")
```

```{r}
ggboxplot(myeloma, x="molecular_group", y="DEPDC1",
          color="molecular_group", add="jitter",
          legend="none") + 
    rotate_x_text(angle = 45) + 
    geom_hline(yintercept = mean(myeloma$DEPDC1),
               linetype=2) + #  add base mean
     stat_compare_means(method = "anova", label.y = 1600)+        # Add global annova p-value
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")                      # Pairwise comparison against all
```

# 8. generate sub figure 

```{r}
p <- ggboxplot(ToothGrowth, x = "supp", y = "len",
          color = "supp",
          add = "point",
          facet.by = "dose", short.panel.labs = FALSE)
# Use only p.format as label. Remove method name.
p + stat_compare_means(label = "p.format")
```

```{r}
# Or use significance symbol as label
p + stat_compare_means(label =  "p.signif", label.x = 1.5)
```

put all figure in one

```{r}
p <- ggboxplot(ToothGrowth, x = "dose", y = "len",
          color = "supp", palette = "jco",
          add = "jitter")
p + stat_compare_means(aes(group = supp))
```

```{r}
# only p value 
p + stat_compare_means(aes(group = supp), label = "p.format")
```

# 9. Other figure 

```{r}
# barplot add mean_se
ggbarplot(ToothGrowth, x = "dose", y = "len", add = "mean_se")+
  stat_compare_means() +                                         # Global p-value
  stat_compare_means(ref.group = "0.5", label = "p.signif",
                     label.y = c(22, 29))                   # compare to ref.group
```

```{r}
ggline(ToothGrowth, x = "dose", y = "len", add = "mean_se")+
  stat_compare_means() +                                         # Global p-value
  stat_compare_means(ref.group = "0.5", label = "p.signif",
                     label.y = c(22, 29))  
```

```{r}
ggbarplot(ToothGrowth, x = "dose", y = "len", add = "mean_se",
          color = "supp", palette = "jco", 
          position = position_dodge(0.8))+
  stat_compare_means(aes(group = supp), label = "p.signif", label.y = 29)
```

```{r}
ggline(ToothGrowth, x = "dose", y = "len", add = "mean_se",
          color = "supp", palette = "jco")+
  stat_compare_means(aes(group = supp), label = "p.signif", 
                     label.y = c(16, 25, 29))
```

## SessionInfo

```{r}
sessionInfo()
```







