---
output: 
  html_document:
    self_contained: false
---


```{r}
library(tidyverse)
```

A functional is a function that takes a function as an input and returns a vector as output. 


```{r}
randomise <- function(f) f(runif(1e3))
randomise(mean)
```

passed mean function to the randomise functipn, which equal

```{r}
mean(runif(1e3))
```


```{r}
library(purrr)

triple <- function(x) x * 3
map(1:3, triple) # map the list to triple function
```

map() defines a mapping from one vector to another. (“Map” also has the nice property of being short, which is useful for such a fundamental building block.)

```{r}
simple_map <- function(x, f, ...) {
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[i]], ...)
  }
  out
}
```

```{r}
# vector produces a vector of the given length and mode.

vector("list", 3) # creat three list
```


```{r}
seq_along(1:3)
```


```{r}
simple_map(1:3, triple)
```

each column in mtcars is a list, so can pass value into map function

```{r}
n_unique <- function(x) length(unique(x))
map_int(mtcars, n_unique) 
```


```{r}
pair <- function(x) c(x, x)
map_dbl(1:2, pair)
```
map function each call to .f must return a single value. If it does not, you’ll get an error:

to fix it:

```{r}
pair <- function(x) (x+x) 
map_dbl(1:2, pair)
```


besides, we also albe to switch back to map(), because map() can accept any type of output. 

```{r}
pair <- function(x) c(x, x)
map(1:2, pair)
```


Anonymous functions and shortcuts

```{r}
map_dbl(mtcars, function(x) length(unique(x)))

# equal to the shortcuts
map_dbl(mtcars, ~ length(unique(.x)))
```



```{r}
as_mapper(~ length(unique(.x)))
```

```{r}
x <- map(1:3, ~ runif(2))
str(x)
```



```{r}
as_mapper(~ runif(2))
```


```{r}
x <- list(
  list(-1, x = 1, y = c(2), z = "a"),
  list(-2, x = 4, y = c(5, 6), z = "b"),
  list(-3, x = 8, y = c(9, 10, 11))
)

x

# Select by name
map_dbl(x, "x")

# Or by both
map_dbl(x, list("y", 1))
```


```{r}
simple_map <- function(x, f, ...) {
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[i]], ...)
  }
  out
}
```




```{r}
boostrap_summary <- function(x, f) {
  f(sample(x, replace = TRUE))
}

simple_map(mtcars, boostrap_summary, f = mean)
```




simple_map() has arguments x and f so you’ll have problems whenever the function you are calling has arguments x or f:



```{r}
simple_map <- function(.x, .f, ...) {
  out <- vector("list", length(.x))
  for (i in seq_along(.x)) {
    out[[i]] <- .f(.x[[i]], ...)
  }
  out
}
```




```{r}
boostrap_summary <- function(x, f) {
  f(sample(x, replace = TRUE))
}

simple_map(mtcars, boostrap_summary, f = mean)
```


```{r}
trims <- c(0, 0.1, 0.2, 0.5)
x <- rcauchy(1000)
```

```{r}
map_dbl(trims, ~ mean(x, trim = .x))

map_dbl(trims, function(trim) mean(x, trim = trim))

# trim: the fraction (0 to 0.5) of observations to be trimmed from each end of x before the mean is computed. Values of trim outside that range are taken as the nearest endpoint.
```

```{r}
mean(x, trim = 0.1)

mean(0.1, x = x)
```

```{r}
by_cyl <- split(mtcars, mtcars$cyl)
by_cyl
```

```{r}
class(by_cyl)

by_cyl$`4`
```

modify(), which always returns the same type of output as the input:

```{r}
df <- data.frame(
  x = 1:3,
  y = 6:4
)

map(df, ~ .x * 2)

modify(df, ~ .x *2)
```


```{r}
xs <- map(1:8, ~ runif(10))
xs[[1]][[1]] <- NA
ws <- map(1:8, ~ rpois(10, 5) + 1)
```

```{r}
xs; ws
```

```{r}
map2_dbl(xs, ws, weighted.mean, na.rm = TRUE)
```


```{r}
x <- map(1:6, ~ sample(1000, 10))
x

imap_chr(x, ~ paste0("The highest value of ", .y, " is ", max(.x)))
```


```{r}
params <- tibble::tribble(
  ~ n, ~ min, ~ max,
   1L,     0,     1,
   2L,    10,   100,
   3L,   100,  1000
)

params
```



```{r}
pmap(params, runif)
```


```{r}
runif(3, 100, 1000)
```


```{r}
x <- c(4, 3, 10)
reduce(x, `+`)

accumulate(x, `+`)
```



```{r}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```


```{r}
f1(1)
```



```{r}
x <- 10
g01 <- function() {
  x <- 20
  x
}

g01()
```


```{r}
i01 <- function(y, z) {
  list(y = y, z = z)
}

i02 <- function(x, ...) {
  i01(...)
}

(i02(x = 1, y = 2, z = 3))
```


```{r}
j04 <- function() invisible(1)

j04()
```



```{r}
j04 <- function() 1

j04()
```



```{r}
(a <- 2)
```


```{r}
a <- b <- c <- d <- 2

a; b; c; d
```


```{r}
j05 <- function() {

  stop("I'm an error")

  return(10)

}

j05()
```


```{r}
j05 <- function() {

  return(10)

}

j05()
```

Replacement functions
Replacement functions act like they modify their arguments in place, and have the special name xxx<-.


```{r}
`second<-` <- function(x, value) {

  x[2] <- value

  x

}

```



```{r}
x <- 1:10

second(x) <- 5L

x
```


```{r}
second(x, 5)
```

```{r}
`modify<-` <- function(x, position, value) {

  x[position] <- value

  x

}

modify(x, 1) <- 10

x

modify(x, 2) <- 10

x
```


```{r}
x <- c(a = 1, b = 2, c = 3)

x

names(x)
```


```{r}
library(rlang) # Functions for Base Types and Core R and 'Tidyverse' Features
# A toolbox for working with base types, core R features like the condition system, and core 'Tidyverse' features like tidy evaluation.
```

The job of an environment is to associate, or bind, a set of names to a set of values.


```{r}
e1 <- env(
  a = FALSE,
  b = "a",
  c = 2.3,
  d = 1:3,
)
```


```{r}
e1
```


```{r}
e1$d <- e1
e1
```

```{r}
env_print(e1)
```

```{r}
env_names(e1)
```



To compare environments, you need to use identical() and not ==. This is because == is a vectorised operator, and environments are not vectors.

```{r}
identical(global_env(), current_env())
```


```{r}
e2a <- env(d = 4, e = 5)
e2b <- env(e2a, a = 1, b = 2, c = 3)
```


```{r}
env_parent(e2b)

env_parent(e2a)
```

```{r}
env_parents(e2b, last = empty_env())
```


```{r}
x <- 0
f <- function() {
  x <<- 1
}
f()
x
```


```{r}
x <- 0
f <- function() {
  x <<- 1
}
# f() if not run this, x will still 0
x
```

can get and set elements of an environment with $ and [[ in the same way as a list:


```{r}
e3 <- env(x = 1, y = 2)
e3$x
```



```{r}
env_get(e3, "x")
```


env_poke()42 takes a name (as string) and a value:

env_bind() allows you to bind multiple values:

```{r}
env_poke(e3, "a", 100)
e3$a

env_bind(e3, a = 10, b = 20)
env_names(e3)

env_unbind(e3, "a")

env_has(e3, "a")
```


```{r}
# 定义teacher对象和行为
teacher <- function(x,...) UseMethod("teacher")
teacher.lecture <- function(x,...) print("上课")
teacher.assignment <- function(x,...) print("布置作业")
teacher.correcting <- function(x,...) print("批改作业")
teacher.default <- function(x,...) print("你不是teacher")

# 定义同学对象和行为

student <- function(x,...) UseMethod("student")
student.attend <- function(x,...) print("听课")
student.homework <- function(x,...) print("写作业")
student.exam <- function(x,...) print("考试")
student.default <- function(x,...) print("你不是student")

#定义两个变量，a老师和b同学

a <- 'teacher' # now, a is still character
class(a)
b <- 'student'

# 给老师变量设置行为

attr(a,"class") <- 'lecture'


class(a)  # now a is a lecture

a
# 执行老师的行为
teacher(a)


attr(a,'class') <- 'assignment'

class(a)

teacher(a)
```



```{r}
where <- function(name, env = caller_env()) {
  if (identical(env, empty_env())) {
    # Base case
    stop("Can't find ", name, call. = FALSE)
  } else if (env_has(env, name)) {
    # Success case
    env
  } else {
    # Recursive case
    where(name, env_parent(env))
  }
}
```



```{r}
where("yyy")
```

```{r}
x <- 5
where("x")

where("mean")
```


package environments

```{r}
search()

search_envs()
```

```{r}
e <- env()
e$g <- function() 1
```

```{r}
# cannot directly get g functions
g()
e$g()
```


Every function in a package is associated with a pair of environments: the package environment, which you learned about earlier, and the namespace environment.


```{r}
sd # so sd in the environment and have a namespace
```



```{r}
library(rlang)
library(ggplot2)
library(scales)

power1 <- function(exp) {
  function(x) {x ^ exp}
}

square <- power1(2)
cube <- power1(3)
```


```{r}
square
```

```{r}
cube
```

how does R find the value associated with exp



```{r}
env_print(square); env_print(cube)
```


```{r}
a <- fn_env(square)
```

```{r}
a$exp
```

```{r}
fn_env(cube)$exp
```


```{r}
power1 <- function(exp) {
  function(x) {x ^ exp}
}

x <- 2
square <- power1(x) # square (x^2), now exp not give to 2

x <- 3 # now x is 3

square(2) # load here, so x is 3 now
```

because x is only evaluated lazily when square() is run, not when power1() is run.



```{r}
power1 <- function(exp) {
  function(x) {x ^ exp}
}

x <- 2
square <- power1(x) # square (x^2)

square(2) # now x is given to exp, and become 2, results is 4

x <- 3

square(2) # still results is 4 

```


```{r}
power2 <- function(exp) {
  force(exp)
  function(x) {
    x ^ exp
  }
}

x <- 2
square <- power2(x)
x <- 3
square(2)
```

```{r}
new_counter <- function() {
  i <- 0
  
  function() {
    i <<- i + 1
    i
  }
}

counter_one <- new_counter()
counter_two <- new_counter()
```

```{r}
counter_one()

counter_one()
```

```{r}
new_counter <- function() {
  i <- 0
  
  function() {
    i <- i + 1
    i
  }
}

counter_one <- new_counter()
counter_two <- new_counter()
```

```{r}
counter_one()

counter_one()
```



```{r}
y <- c(12345, 123456, 1234567)

comma_format()(y)

comma_format(y)
```

```{r}
comma_format 
```


```{r}
a <- comma_format() # a is a function now

a(y)
```


```{r}
df <- data.frame(x = 1, y = y)

core <- ggplot(df, aes(x, y)) + 
  geom_point() + 
  scale_x_continuous(breaks = 1, labels = NULL) +
  labs(x = NULL, y = NULL)
  
core

core + scale_y_continuous(
  labels = comma_format()
)

core + scale_y_continuous(
  labels = number_format(scale = 1e-3, suffix = " K")
)

core + scale_y_continuous(
  labels = scientific_format()
)
```


```{r}
# construct some sample data with very different numbers in each cell
sd <- c(1, 5, 15)
n <- 100

df <- data.frame(x = rnorm(3 * n, sd = sd), sd = rep(sd, n))

ggplot(df, aes(x)) + 
  geom_histogram(binwidth = 2) + 
  facet_wrap(~ sd, scales = "free_x") + 
  labs(x = NULL)
```


```{r}
binwidth_bins <- function(n) {
  force(n)
  
  function(x) {
    (max(x) - min(x)) / n
  }
}

ggplot(df, aes(x)) + 
  geom_histogram(binwidth = binwidth_bins(20)) + 
  facet_wrap(~ sd, scales = "free_x") + 
  labs(x = NULL)
```


```{r}
a <- fn_env(binwidth_bins(20))

a$n
```


```{r}
names <- list(
  square = 2, 
  cube = 3, 
  root = 1/2, 
  cuberoot = 1/3, 
  reciprocal = -1
)
funs <- purrr::map(names, power1)
```

```{r}
funs$square(64)
```

```{r}
with(funs, root(100))
# with(data, expr, ...)
# Evaluate an R expression in an environment constructed from data, possibly modifying (a copy of) the original data.

with(mtcars, mpg[cyl == 8  &  disp > 350])
```


For a longer effect, you can attach() the functions to the search path, then detach() when you’re done:


```{r}
attach(funs) 
# attach This means that the database is searched by R when evaluating a variable, so objects in the database can be accessed by simply giving their names.
root(100)
```

A function operator is a function that takes one (or more) functions as input and returns a function as output.

If you’re familiar with Python, decorators is just another name for function operators.

装饰器(Decorators)是Python 的一个重要部分。 简单地说：他们是修改其他函数的功能的函数。 他们有助于让我们的代码更简短，也更Pythonic（Python范儿）。

```{r}
chatty <- function(f) {
  force(f)
  
  function(x, ...) {
    res <- f(x, ...)
    cat("Processing ", x, "\n", sep = "")
    res
  }
} ## return a function


f <- function(x) x ^ 2 

s <- c(3, 2, 1)

purrr::map_dbl(s, chatty(f)) # map list to the functions
```


```{r}
f
```


```{r}
chatty(f)
```


```{r}
library(purrr)
library(memoise)
```


```{r}
x <- list(
  c(0.512, 0.165, 0.717),
  c(0.064, 0.781, 0.427),
  c(0.890, 0.785, 0.495),
  "oops"
)

str(x)

out <- rep(NA_real_, length(x)) # There are also constants NA_integer_, NA_real_, NA_complex_ and NA_character_ of the other atomic vector types which support missing values:

out

for (i in seq_along(x)) {
  out[[i]] <- sum(x[[i]])
}

out
```

```{r}
map_dbl(x, sum)
```

safely() is a function operator that transforms a function to turn errors into data. 


```{r}
sum
```


```{r}
safe_sum <- safely(sum)
safe_sum
```


```{r}
x
out <- map(x, safely(sum))
out
```

```{r}
out <- transpose(map(x, safely(sum))) # Transpose turns a list-of-lists "inside-out"; it turns a pair of lists into a list of pairs, or a list of pairs into pair of lists.
out$result

str(out)

# str: Compactly display the internal structure of an R object, a diagnostic function and an alternative to summary 
```

R语言一个重点是面向对象编程（object-oriented programming）
Python从设计之初就已经是一门面向对象的语言，正因为如此，在Python中创建一个类和对象是很容易的。
所以可以同时在R和python中学面对对象编程。

OO systems call the type of an object its class, and an implementation for a specific class is called a method. Roughly speaking, a class defines what an object is and methods describe what that object can do. 

object.method(arg1, arg2)

generic(object, arg2, arg3)

You might wonder if S1 and S2 exist. They don’t: S3 and S4 were named according to the versions of S that they accompanied. The first two versions of S didn’t have any OOP framework.

sloop::otype(). It makes it easy to figure out the OOP system used by a wild-caught object:

```{r}
library(sloop)

mle_obj <- stats4::mle(function(x = 1) (x - 2) ^ 2)
otype(mle_obj)
```



```{r}
sloop::otype(1:10)

sloop::otype(mtcars)
```

Technically, the difference between base and OO objects is that OO objects have a “class” attribute:

```{r}
attr(1:10, "class")
#> NULL

attr(mtcars, "class")
#> [1] "data.frame"

# attr: Get or set specific attributes of an object.
```


only OO objects have a class attribute, every object has a base type:

```{r}
typeof(1:10)
#> [1] "integer"

typeof(mtcars)
#> [1] "list"
```


```{r}
f <- factor(c("a", "b", "c"))

typeof(f)
```

```{r}
attributes(f)
```

Many base R functions are generic, including the important print():


```{r}
ftype(print)
```


generic finds that method by performing method dispatch (方法调度).


```{r}
s3_dispatch(print(f))
```

```{r}
print.factor(f)
```

use sloop::s3_get_method(), which will work regardless of where the method lives:

```{r}
s3_get_method(t.data.frame) # only work for method
```


```{r}
# Create and assign class in one step
x <- structure(list(), class = "my_class")

# structure returns the given object with further attributes set.

# Create, then set class
x <- list()
class(x) <- "my_class"
```

```{r}
class(x)
```

```{r}
mean
```

```{r}
my_new_generic <- function(x) {
  UseMethod("my_new_generic")
}
```

```{r}
my_new_generic
```

```{r}
x <- Sys.Date()

print(x)

s3_dispatch(print(x))
```

The output here is simple:

=> indicates the method that is called, here print.Date()
* indicates a method that is defined, but not called, here print.default().

```{r}
s3_methods_generic("mean")

s3_methods_generic("print")
```


# find all methods defined for a generic or associated with a class

```{r}
s3_methods_class("ordered")
```




```{r}
x <- list()
class(x) <- "my_class"

x
```


```{r}
teacher <- function(x,...) UseMethod("teacher") # 这时teacher也变成了一个对象

teacher
```

```{r}
teacher.lecture <- function(x,...) print("讲课")
teacher.assignment <- function(x,...) print("布置作业")
teacher.correcting <- function(x,...) print("批改作业")
teacher.default <- function(x,...) print("你不是teacher")
```

```{r}
teacher
```

```{r}
a <- "teacher" # 这时a是个字符串

a
```

```{r}
class(a)
```



```{r}
attr(a,"class") <- 'lecture'
```

```{r}
a
```

```{r}
attributes(a)
```


```{r}
## 第一步：定义一个泛型函数，这个函数的函数体只有一个固定的语句
doit = function(...) {
  xx = list(...)
  UseMethod("doit", xx[[1]], xx[[2]])
  }

## 第二步：设置针对特定CLASS的动作函数
doit.character = function(...) {
    cat("With STRING class attribute\n")
}
doit.integer = function(...) {
    cat("With INTEGER class attribute\n")
}
doit.God = function(...) {
    cat("With God class attribute\n")
}
### 第三步：设置一个默认的动作函数
doit.default = function(...) {
    cat("UNKNOWN class attribute\n")
}

doit(1:10, "abc")


##上面三个步骤就完成了doit泛型函数的设置。看看效果吧：
a = "ABCDE"

doit(a) # 这里a本来是string
## With STRING class attribute

attr(a, "class") = "integer"
class(a) #[1] "integer"
doit(a)
## With INTEGER class attribute

attr(a, "class") = "God"
doit(a)
## With God class attribute

a = as.factor(a)
doit(a)
## UNKNOWN class attribute
```


```{r}
# 用UseMethod()定义teacher泛型函数
teacher <- function(x,...) UseMethod("teacher") # 这时teacher也变成了一个对象
# 用pryr包中ftype()函数，检查teacher类型
ftype(teacher)

# 定义teacher内部函数, 都是teacher的属性
teacher.lecture <- function(x,...) print("讲课")
teacher.assignment <- function(x,...) print("布置作业")
teacher.correcting <- function(x,...) print("批改作业")
teacher.default <- function(x,...) print("你不是teacher")
teacher.character <- function(x, ...) print("请先定义")
```


```{r}
a <- "teacher" # 这时a是个字符串
teacher(a)

# 给老师变量设置行为
attr(a,"class") <- 'lecture' # 给a一个类属性 lecture


attr(a,"class") <- c('correcting', 'lecture') # 给a一个类属性 lecture

# 执行老师的行为
teacher(a)
teacher(a)



attributes(a)

# $class
# [1] "lecture"
```



```{r}
# 用UseMethod()定义teacher泛型函数
teacher <- function(...) {
  xx = list(...)
  UseMethod("teacher", xx[[1]], xx[[2]])} # 这时teacher也变成了一个对象
# 用pryr包中ftype()函数，检查teacher类型
ftype(teacher)

# 定义teacher内部函数, 都是teacher的属性
teacher.lecture <- function(x,...) print("讲课")
teacher.assignment <- function(x,...) print("布置作业")
teacher.correcting <- function(x,...) print("批改作业")
teacher.default <- function(x,...) print("你不是teacher")
teacher.character <- function(x, ...) print("请先定义")

a <- "teacher" # 这时a是个字符串
attr(a,"class") <- c('correcting', 'lecture') # 给a一个类属性 lecture
teacher(a)

```


```{r}
doit = function(...) UseMethod("doit")
doit.character = function(...) {
    cat("With STRING class attribute\n")
    NextMethod()
}

doit.integer = function(...) {
    cat("With INTEGER class attribute\n")
    NextMethod()
}

doit.God = function(...) {
    cat("With God class attribute\n")
    NextMethod()
}

doit.default = function(...) {
    cat("UNKNOWN class attribute\n")
}

## 多CLASS属性对象
x = "abc"
class(x) = c("UNKNOWN", "integer", "character", "God")
doit(x)
# With INTEGER class attribute
# With STRING class attribute
# With God class attribute
# UNKNOWN class attribute
```

```{r}
x = "abcd"

attr(x, "class") <- c("UNKNOWN", "integer", "character", "God")

doit(x)
```



```{r}
j <- list(name = "Joe", salary=5000, union=T)

class(j) <- "employee"

print.employee <- function(wrkr){
cat("Name:", wrkr$name, "\n")
cat("Salary: ", wrkr$salary, "\n")
cat("Union member: ", wrkr$union, "\n")
}

print(j)


# Name: Joe 
# Salary:  5000 
# Union member:  TRUE 
```



```{r}
# 创建类
my_person <- function(ln, a, ht){
  structure(list(lastName = ln, age = a, height = ht), class = "my_person")
}

## 创建第一个泛型函数
display <- function(obj) UseMethod("display")
# 创建方法
display.my_person <- function(obj){
  cat("Last name : ", obj$lastName, "\n")
  cat("Age       : ", obj$age, "\n")
  cat("Height    : ", obj$height, "\n")
}


## 创建第二个泛型函数，如果泛型函数的方法又多个参数，务必加上...
nYear <- function(obj,...) UseMethod("nYear")
# 创建方法
nYear.my_person <- function(obj,n){
  obj$age <- obj$age + n
  cat("Age after",n,"years is:", obj$age, "\n")
}

# 由类创建对象
Tom <- my_person("Godden", 19, 176)

class(Tom)

# 第一个泛型函数派送
display(Tom)
## Last name :  Godden 
## Age       :  19 
## Height    :  176

# 第二个泛型函数派送
nYear(Tom,7)
## Age after 7 years is: 26
```

## S4 for programming

```{r}
# 创建Class - Person
setClass("Person", slots = list(name = "character", age = "numeric"))

# 创建Class - Employee，利用contains表明参数继承自Person。
setClass("Employee", slots = list(boss = "Person"), contains = "Person")

# 设置基于Class - Person的对象
alice <- new("Person", name = "Alice", age = 40)

# 设置基于Class - Employee的对象
john <- new("Employee", name = "John", age = 20, boss = alice)


str(alice)
str(john) # 可以看到jhon多了一个slot，即@boss
```


```{r}
# Formal class 'Employee' [package ".GlobalEnv"] with 3 slots
#   ..@ boss:Formal class 'Person' [package ".GlobalEnv"] with 2 slots
#   .. .. ..@ name: chr "Alice"
#   .. .. ..@ age : num 40
#   ..@ name: chr "John"
#   ..@ age : num 20
```


```{r}
setClass("Person", slots = list(name = "character", age = "numeric"), prototype = list(age = 20))

setValidity("Person",  
 function(object){
 if (object@age < 18)
 "he/she is too yong to get the job."
 })

new("Person", name = "Penny", age = 17)

# Error in validObject(.Object) : invalid class “Person” object: he/she is too yong to get the job.
```


```{r}
# 查看alice对象的name属性
alice@name

slot(alice, "age")

alice@age
```


```{r}
getClasses()
```


```{r}
work<-function(x) cat(x, "is working")
work('Conan')
```


```{r}
# S4的方法
# 定义Person对象
setClass("Person",slots=list(name="character",age="numeric"))
# 定义泛型函数work，即接口
setGeneric("work",function(object) standardGeneric("work"))
# 定义work的现实，并指定参数类型为Person对象
setMethod("work", signature(object = "Person"), function(object) cat(object@name , "is working") )
# 创建一个Person对象a
a<-new("Person",name="Conan",age=16)
# 把对象a传入work函数
work(a)
# Conan is working
```

```{r}
ftype(work)
```



```{r}

```






















