---
title: Functional programming-Map, Reduce, Filter and Lambda
author: Liang
date: '2018-06-17'
slug: functional-programming-map-reduce-filter-and-lambda
categories:
  - R
tags:
  - R
---

This post will talk about four useful function in R programming. Map, Reduce, Filter and Lambda.

# LAMBDA

Lambda can be seen as a short (normally one line) function definition. There is no particular syntax for lambda in R, except that you don’t need to assign the function to a variable (function name), and also you will need to enclose the function using brackets or curly brace. For example:

```
(function add(x, y) x+y)
{function add(x, y) x+y}
```

We can pass the lambda function as a parameter to other functions e.g. Map, Reduce or Filter.

# Map

Map function allows the mapping from one vector to another using a map function, which can be specified by lambda. For example, let’s define a vector from 1 to 100.

```{r}
x = 1:100
```

If we want to convert each element of the vector to its doubles, so we can pass a lambda to the Map function, like this:

```{r}
y=Map({function (a) a*2}, x)
```

However, the variable y will store a list of vectors instead of a single vector. We can use the unlist function to unroll the list of vectors into a single vector.

```{r}
unlist(y)
```

# Reduce 

Reduce will perform the function on a list of vectors one by one, and finally return a single value.

```{r}
x=seq(1,10,0.5)
Reduce({function (x, y) x+y}, x)
```

The lambda function can be instead with `sum` here.
```{r}
Reduce(sum, x)
```

# Filter

The function Filter will remove all elements when they do not satisfy the condition (function returns false)

```{r}
x = 1:10
Filter({function (x) x%%2==0}, x)
```







