---
title: ggplot2 plotting with error bars
author: Liang
date: '2018-05-03'
slug: ggplot2-plotting-with-error-bars
categories:
  - R
tags:
  - R
---
## Problem

As a biologist, we need to plot a lot of data every day. To make our result more convincing, usually we need to add error bars on the bar plot results. In this Poster, I will show how to use ggplot2 to make a bar plot with error bar.

## Solution

### Prepare Data

The data we are using in this practice is `ToothGrowth`, we can get the data in `datasets` library. first we can look at the dataset.
```{r}
library(datasets)
library(ggplot2)

data(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)
```

Next, We will use `summarySE` function to summary the dataset. Please consider about the difference between `summary` and `summarySE`. `summarySE` is in `Rmisc` library, so first require `Rmisc`

```{r, warning=FALSE, message=FALSE}
library(Rmisc)
tg <- ToothGrowth
# summarySE provides the standard deviation, standard error of the mean, and a (default 95%) confidence interval
tgc <- summarySE(tg, measurevar="len", groupvars=c("supp","dose"))
tgc
#>   supp dose  N   len       sd        se       ci
#> 1   OJ  0.5 10 13.23 4.459709 1.4102837 3.190283
#> 2   OJ  1.0 10 22.70 3.910953 1.2367520 2.797727
#> 3   OJ  2.0 10 26.06 2.655058 0.8396031 1.899314
#> 4   VC  0.5 10  7.98 2.746634 0.8685620 1.964824
#> 5   VC  1.0 10 16.77 2.515309 0.7954104 1.799343
#> 6   VC  2.0 10 26.14 4.797731 1.5171757 3.432090
```

After the data is summarized, we can make the graph. I will show the basic bar graph in ggplot2 with error bars representing either the standard error of the mean, or 95% confidence interval.

### Bar plot with ggplot

```{r}
# Use dose as a factor rather than numeric
tgc$dose <- factor(tgc$dose)

ggplot(tgc, aes(x=dose, y=len, fill=supp)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black", # Use black outlines,
             size=.3) +      # Thinner lines
    geom_errorbar(aes(ymin=len-se, ymax=len+se),
                  size=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) + # adjust errorbar`s position
    xlab("Dose (mg)") +
    ylab("Tooth length") +
    scale_fill_hue(name="Supplement type", # Legend label, use darker colors
                   breaks=c("OJ", "VC"),
                   labels=c("Orange juice", "Ascorbic acid")) +
    ggtitle("The Effect of Vitamin C on\nTooth Growth in Guinea Pigs") +
    scale_y_continuous(breaks=0:20*4) +
    theme_classic()
```

`scale_fill_hue` is one way to modify the legend. And also there some other way in ggplot2.


Reference:
[Plotting means and error bars (ggplot2)](http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/)