---
title: "Meachine learning in R - Supervised Learning"
author: "KJY"
date: '2022-06-07'
slug: mechine_learning_in_R2
categories: R
tags: []
lastmod: '2022-06-07T10:29:22-05:00'
keywords: []
description: ''
comment: no
toc: no
autoCollapseToc: no
postMetaInFooter: no
hiddenFromHomePage: no
contentCopyright: no
reward: no
mathjax: yes
mathjaxEnableSingleDollar: yes
mathjaxEnableAutoNumber: yes
hideHeaderAndFooter: no
flowchartDiagrams:
  enable: no
  options: ''
sequenceDiagrams:
  enable: no
  options: ''
---

## KNN 

K-nearest neighbours works by directly measuring the (Euclidean) distance between observations and inferring the class of unlabelled data from the class of its nearest neighbours.


Typically in machine learning, there are two clear steps, where one first trains a model and then uses the model to predict new outputs (class labels in this case). In the kNN, these two steps are combined into a single function call to knn.

```{r}
set.seed(12L)

# from 150 numbers, randomly select 50 numbers
tr <- sample(150, 50)
nw <- sample(150, 50)


library("class")
knnres <- knn(iris[tr, -5], iris[nw, -5], iris$Species[tr])

knnres

head(knnres)
```


```{r}
table(knnres, iris$Species[nw])
```


```{r}
library(magrittr)
knnres == iris$Species[nw] %T>% mean()
```


```{r}
mean(knnres == iris$Species[nw])
```


parameter k defines how many nearest neighbours will be considered to assign a class to a new unlabelled observation


```{r}
args(knn)
```


the default value is 1

## Model performance

```{r}
library(caret)
```


```{r}
data(diamonds)

model <- lm(price ~ ., diamonds)

model
# predict is a generic function for predictions from the results of various model fitting functions.
p <- predict(model, diamonds)
```

```{r}
## Error on prediction
error <- p - diamonds$price
rmse_in <- sqrt(mean(error^2)) ## in-sample RMSE root mean squared error (RMSE)
rmse_in
```

```{r}
set.seed(42)
ntest <- nrow(diamonds) * 0.80
test <- sample(nrow(diamonds), ntest)
model <- lm(price ~ ., data = diamonds[test, ])
p <- predict(model, diamonds[-test, ])
error <- p - diamonds$price[-test]
rmse_out <- sqrt(mean(error^2)) ## out-of-sample RMSE
rmse_out
```































