---
title: "One sample proportion test (单比率检验)"
date: '2022-05-28'
author: KJY
slug: One_sample_proportion
draft: false
categories:
  - R
tags: []
lastmod: '2022-05-28T10:30:09-07:00'
keywords: []
description: ''
comment: no
toc: no
autoCollapseToc: no
postMetaInFooter: no
hiddenFromHomePage: no
contentCopyright: no
reward: no
mathjax: yes
mathjaxEnableSingleDollar: yes
mathjaxEnableAutoNumber: yes
hideHeaderAndFooter: no
flowchartDiagrams:
  enable: no
  options: ''
sequenceDiagrams:
  enable: no
  options: ''
---

---
title: "One sample proportion test (单比率检验)"
date: 2022-05-28T16:03:40-06:00
author: KJY
slug: One_sample_proportion
draft: false
toc: true
categories:  
  -statis
tags:        
  - article
---

Reference:


https://www.r-bloggers.com/2022/05/one-sample-proportion-test-in-r-complete-guide/

https://zhuanlan.zhihu.com/p/349916370

One sample proportion test in R, when there are just two categories, the one proportion Z-test is used to compare an observed proportion to a theoretical one.

have a population that is half male and half female (p = 0.5 = 50%). Some of these total (n = 160), including 100 males and 60 females, acquired a spontaneous malignancy.

[Malignancy (from Latin male 'badly', and -gnus 'born') is the tendency of a medical condition to become progressively worse.]

Question: We’d like to know if cancer affects more men than women.

The observed male proportion (po) is 100/160.
The observed female percentage (q) is 1-po.
The predicted male proportion (pe) is 0.5. (50 percent )
A total of 160 observations (n) were made.

The following are the relevant alternative hypothesis (H1)

H1:po≠pe (different); hypotheses 1
H1:po>pe (greater); hypotheses 2
H1:po<pe (less); hypotheses 3
Note that:

Two-tailed tests are used to test hypotheses 1.
One-tailed tests are used to test hypotheses 2 and 3.

To do a one-proportion test, use the R methods binom.test() and prop.test():

Calculate the exact binomial test with binom.test(). When the sample size is small, prop.test() is recommended.



```
binom.test(x, n, p = 0.5, alternative = "two.sided")
prop.test(x, n, p = NULL, alternative = "two.sided", correct = TRUE)
```

x: the number of successes
n: the total number of trials
p: the probability to test against.
correct: a logical indicator of whether Yates’ continuity correction should be used if at all practicable.

```{r}
prop <- prop.test(x = 100, n = 160, p = 0.5, correct = FALSE)
prop
```


If you wish to see if the percentage of men with cancer is less than 0.5 (one-tailed test), enter:

```{r}
prop.test(x = 100, n = 160, p = 0.5, correct = FALSE, alternative = "less")
```


# Other information

单一变量的假设检验是对数据中某一个变量（数值变量、或多项有序分类变量、或多项无序分类变量、或二项分类变量）所进行的统计分析，常常比较该样本统计量与总体参数或中位数之间是否不同。

> 一个变量和总体之间的比较

例如下问就是很标准的单样本检测的问题

某校18岁男生的身高情况如表1，已知全国18岁男性身高为172cm，问该校18岁男生的身高是否与全国水平相同？

或者是

产线设定的产品不良率是99%，然后我们收集1000的产品数据，不良数是999.查看不良率是否等于我们假定的99%？


A Z-score is a numerical measurement that describes a value's relationship to the mean of a group of values. Z-score is measured in terms of standard deviations from the mean. If a Z-score is 0, it indicates that the data point's score is identical to the mean score.

A z-table, also called the standard normal table, is a mathematical table that allows us to know the percentage of values below (to the left) a z-score in a standard normal distribution (SND)











